### 2023-03-23 main.tex
ii
FixIntegrability
bibFix

\vspace{.5cm}
\noindent
{\bf Remark:} in the following we will often consider type $1$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types (at the same site). It is interesting to observe that, in case we stop distinguishing the colors of the ``true'' particles of types $\{2,\ldots,N\}$, we retrieve the 
boundary-driven version of the partial symmetric exclusion process \cite{ss,carinci2013duality}.
The choice of type $1$ for the distinguished type is arbitrary and does not affect
the results. In particular, we shall see in Section \ref{sectionDuality} about duality, that
the dual process eventually has only particles of type $1$ on the graph.
Interpreting type $1$ as an hole, one can then say that the dual process voids the graph.

@@

    Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
[E,E_{1\ell}]=\sum_{a=2}^{N}\left(E_{a\ell}\delta_{11}-E_{11}\delta_{a\ell}\right)=\sum_{j=2}^{N}E_{j\ell}-E_{11};
\end{equation}
and 
\begin{equation}
\begin{split}
\left[E,[E,E_{1\ell}]\right]&=\sum_{b=2}^{N}\sum_{j=2}^{N}\left[E_{b1},E_{j\ell}\right]-\sum_{b=2}^{N}\left[E_{b1},E_{11}\right]
\\&=
\sum_{j,b=2}^{N}\left(E_{b\ell}\delta_{j1}-E_{j1}\delta_{b\ell}\right)-\sum_{b=2}^{N}\left(E_{b1}\delta_{11}-E_{11}\delta_{b1}\right)
\\=&
-2\sum_{j=2}^{N}E_{j1};
\end{split}
\end{equation}
Inserting the above commutators in \eqref{HadamardFormula}, we obtain \eqref{HT-E1l}.
\item for $A=E_{k\ell}$ with $k,\ell \in \{2,\ldots,N\}$ we obtain 
\begin{equation}\label{HT-Ekl}
    \exp{(-E)}E_{\ell k}\exp{(E)}=E_{k\ell}+E_{\ell 1}
\end{equation}
\com{double check}
   Indeed, using \eqref{eq:comgl} we have 
   {\color{red}
\begin{equation}
[E,E_{k\ell}]=\sum_{a=2}^{N}\left(E_{ak}\delta_{\ell 1}-E_{\ell 1}\delta_{ak}\right)=-E_{\ell 1};
\end{equation}
}
\begin{equation}
[E,E_{k\ell}]=\sum_{a=2}^{N}\left(E_{a\ell}\delta_{k1}-E_{k1}\delta_{\ell a}\right)=-E_{k1};
\end{equation}
Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT-Ekl}.
\item for $A=E_{11}$ we obtain 
\begin{equation}\label{HT-E11}
    \exp{(-E)}E_{11}\exp{(E)}=E_{11}-\sum_{j=2}^{N}E_{j1}
\end{equation}
  Indeed, using \eqref{eq:comgl} we have

@@

We aim to find an explicit for for the steady state $|\Psi\rangle$. The MPA states the form of the steady state in function of an abstract algebra of operators. If an explicit representation of $(X_{a})_{a=1,\ldots,N}$ was available, it would be possible to compute explicitly all the correlations and currents of particles for the Hamiltonian \eqref{hamiltonian}. Another way of computing this steady state, without passing through a representation, is using the commutators \eqref{bulk} and the actions \eqref{leftBoundary} and \eqref{rightBoundary}. For the one and two point correlation and for the currents this has been done in \cite{vanicat}. Inspired by this last reference, in this subsection, we aim to find exact formulas for the steady state of \eqref{hamiltonian} using only the \eqref{bulk}, \eqref{leftBoundary} and \eqref{rightBoundary}. In subsection \eqref{correlation-section}, this result will leads to exact correlation for arbitrary number of particles of type $\{2,\ldots,N\}$. This aim will be achieved by the help of a similarity transformation $\mathcal{S}$ that makes commutation relation simpler, and allows to find the steady state of a transformed Hamiltonian, that will be denoted by $\mathbb{H}$, with lower-triangular left boundary and diagonal right boundary. Therefore, inverting this similarity, one can retrieve the steady state of \eqref{hamiltonian}. A fundamental 'ingredient' of these computation is duality, that has been proved in section~\ref{sectionDuality}. \com{clarify}

@@

    S:=\sum_{a=2}^{N}\beta_{a}\sum_{b=2}^{N}e_{ab}+V=-\sum_{a=2}^{N}\beta_{a}\sum_{b=2}^{N}e_{ab}+\exp{\left(\sum_{a=2}^{N}e_{1a}\right)}

@@

			\widehat{\Psi}(\bm{s})=\sum_{i=1}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{i},1}}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(L+2)}\prod_{l=1}^{L}\left(\lambda_{s_{l}}\left(1+L-l-\sum_{j=l}^{L}c_{j}\right)\right)^{c_{l}}\beta_{s_{l}}^{(1-c_{l})(1-\delta_{s_{l},1})}

@@

	\left(T^{T}\otimes T^{T}\right)_{s_{1},s_{2}}=e_{s_{1},s_{2}}^{s_{1},s_{2}}+(1-\delta_{s_{2},1})\beta_{s_{2}}e_{s_{1},s_{2}}^{s_{1},1}+(1-\delta_{s_{1},1})\beta_{s_{1}}e_{s_{1},s_{2}}^{1,s_{2}}+(1-\delta_{s_{1},1})(1-e_{s_{2},1})\beta_{s_{1}}\beta_{s_{2}}e_{s_{1},s_{2}}^{1,1}

@@

  \left(1+\sum_{j=i}^{L}\delta_{s_{j},1}\right) =\left(1+L-i-\sum_{j=i}^{L}(1-\delta_{s_{i},1})\right)

@@

	\widetilde{\Psi}(\bm{s})=\frac{\Gamma(2+\sum_{i=1}^{L}\delta_{s_{i},1})}{\Gamma(L+2)}\prod_{i=1}^{L}\left(\lambda_{s_{i}}\left(1+L-i-\sum_{j=i}^{L}(1-\delta_{s_{i},1})\right)\right)^{1-\delta_{s_{i},1}}

@@

		\widehat{\Psi}(\mathbf{s})=\sum_{c_{1}=0}^{1-\delta_{s_{1},1}}\ldots\sum_{c_{L}=0}^{1-\delta_{s_{L},1}}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(L+2)}\prod_{i=1}^{L}\left[\lambda_{s_{i}}\left(1+L-i-\sum_{j=i}^{L}c_{j}\right)\right]^{c_{i}}\left(\beta_{s_{i}}\right)^{(1-c_{i})(1-\delta_{s_{i},1})}
Indeed, when $c_{i}=0$ this formula replaces the contribution in the square brackets with a $b_{s_{i}}$. 

@@

        \Psi(\bm{s})&=\sum_{i=1}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{i},1}}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(L+2)}\prod_{l=1}^{L}\left(\lambda_{s_{l}}\left(1+L-l-\sum_{j=l}^{L}c_{j}\right)\right)^{c_{l}}\beta_{s_{l}}^{(1-c_{l})(1-\delta_{s_{l},1})}

@@

        \sum_{r=1}^{\ell}\sum_{d_{q_{r}}=2}^{N}\sum_{i=1}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{i},1}}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(2+L)}

@@

  \prod_{i=1\,:\, i\neq q_{1},\ldots,q_{\ell}}^{L}\left(\lambda_{s_{i}}\left(1-L-i-\sum_{j=i}^{L}c_{j}\right)\right)^{c_{i}}\beta_{s_{i}}^{(1-c_{i})(1-\delta_{s_{i},1})}
  \prod_{r=1}^{\ell}\lambda_{q_{r}}\left(1+L-q_{r}-\sum_{j=q_{r}}^{L}c_{j}\right)

@@

  &\widehat{\Psi}(\ldots,a_{i},\ldots,d_{q_{r}},\ldots)=\sum_{i=1}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{i},1}}\left(\prod_{r=1}^{\ell}\delta_{c_{r},1}\right)\frac{\Gamma(2+L-\sum_{k=1}^{L}}{\Gamma(2+L)}

@@

  \prod_{i=1\,:\, i\neq q_{1},\ldots,q_{\ell}}^{L}\left(\lambda_{s_{i}}\left(1-L-i-\sum_{j=i}^{L}c_{j}\right)\right)^{c_{i}}\beta_{s_{i}}^{(1-c_{i})(1-\delta_{s_{i},1})}
  \prod_{r=1}^{\ell}\lambda_{q_{r}}\left(1+L-q_{r}-\sum_{j=q_{r}}^{L}c_{j}\right)
indeed, in coordinates $q_{r}$ with $r=1,\ldots \ell$ the site is non-empty, thus $c_{q_{r}}=1$. Finally, considering all possible occupations $\bm{s}=(s_{1},\ldots,s_{L})$ we obtain \eqref{steadyStateH}. 

@@

	\widehat{\Psi}_{\bm{y}}(s_{1},\ldots,s_{m})=\sum_{i=1}^{L}\sum_{c_{i}=0}^{1}\frac{\Gamma(2+L-\sum_{k=1}^{m}c_{k})}{\Gamma(2+L)}\prod_{k=1}^{m}\lambda_{s_{k}}^{c_{k}}\beta_{s_{k}}^{1-c_{k}}\left(L+1-y_{k}-\sum_{j=k}^{m}c_{j}\right)^{c_{k}}

>>>>

\vspace{.5cm}
\noindent
{\bf Remark:} in the following we will often consider type $1$ as a distinguished type
(called `holes'), whose number at each site is fixed by the numbers of particles 
of the remaining $N-1$ types (at the same site). It is interesting to observe that, in case we stop distinguishing the colors of the ``true'' particles of types $\{2,\ldots,N\}$, we retrieve the 
boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
The choice of type $1$ for the distinguished type is arbitrary and does not affect
the results. In particular, we shall see in Section \ref{sectionDuality} about duality, that
the dual process eventually has only particles of type $1$ on the graph.
Interpreting type $1$ as an hole, one can then say that the dual process voids the graph.

@@

    Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
[E,E_{1\ell}]=\sum_{a=2}^{N}\left(E_{a\ell}\delta_{11}-E_{11}\delta_{a\ell}\right)=\sum_{j=2}^{N}E_{j\ell}-E_{11};
\end{equation}
and 
\begin{equation}
\begin{split}
\left[E,[E,E_{1\ell}]\right]&=\sum_{b=2}^{N}\sum_{j=2}^{N}\left[E_{b1},E_{j\ell}\right]-\sum_{b=2}^{N}\left[E_{b1},E_{11}\right]
\\&=
\sum_{j,b=2}^{N}\left(E_{b\ell}\delta_{j1}-E_{j1}\delta_{b\ell}\right)-\sum_{b=2}^{N}\left(E_{b1}\delta_{11}-E_{11}\delta_{b1}\right)
\\=&
-2\sum_{j=2}^{N}E_{j1};
\end{split}
\end{equation}
Inserting the above commutators in \eqref{HadamardFormula}, we obtain \eqref{HT-E1l}.
\item for $A=E_{k\ell}$ with $k,\ell \in \{2,\ldots,N\}$ we obtain 
\begin{equation}\label{HT-Ekl}
    \exp{(-E)}E_{\ell k}\exp{(E)}=E_{k\ell}+E_{\ell 1}
\end{equation}
   Indeed, using \eqref{eq:comgl} we have 
\begin{equation}
[E,E_{\ell k}]=\sum_{a=2}^{N}\left(E_{ak}\delta_{\ell 1}-E_{\ell 1}\delta_{ak}\right)=-E_{\ell 1};
\end{equation}
\begin{equation}
[E,E_{k\ell}]=\sum_{a=2}^{N}\left(E_{a\ell}\delta_{k1}-E_{k1}\delta_{\ell a}\right)=-E_{k1};
\end{equation}
Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT-Ekl}.
\item for $A=E_{11}$ we obtain 
\begin{equation}\label{HT-E11}
    \exp{(-E)}E_{11}\exp{(E)}=E_{11}-\sum_{j=2}^{N}E_{j1}
\end{equation}
  Indeed, using \eqref{eq:comgl} we have

@@

We aim to find an explicit for for the steady state $|\Psi\rangle$. The MPA states the form of the steady state in function of an abstract algebra of operators. If an explicit representation of $(X_{a})_{a=1,\ldots,N}$ was available, it would be possible to compute explicitly all the correlations and currents of particles for the Hamiltonian \eqref{hamiltonian}. Another way of computing this steady state, without passing through a representation, is using the commutators \eqref{bulk} and the actions \eqref{leftBoundary} and \eqref{rightBoundary}. For the one and two point correlation and for the currents this has been done in \cite{vanicat2017exact}. Inspired by this last reference, in this subsection, we aim to find exact formulas for the steady state of \eqref{hamiltonian} using only the \eqref{bulk}, \eqref{leftBoundary} and \eqref{rightBoundary}. In subsection \eqref{correlation-section}, this result will leads to exact correlation for arbitrary number of particles of type $\{2,\ldots,N\}$. This aim will be achieved by the help of a similarity transformation $\mathcal{S}$ that makes commutation relation simpler, and allows to find the steady state of a transformed Hamiltonian, that will be denoted by $\mathbb{H}$, with lower-triangular left boundary and diagonal right boundary. Therefore, inverting this similarity, one can retrieve the steady state of \eqref{hamiltonian}. A fundamental 'ingredient' of these computation is duality, that has been proved in section~\ref{sectionDuality}. Indeed, the similarity transformation $\mathcal{S}$ relies on an "intermediate transformation" $V$ (see \eqref{transformationV}) that is, up to the diagonal operator $R$, the duality matrix. \com{clarify}

@@

    S:=\sum_{a=2}^{N}\beta_{a}\sum_{b=1}^{N}e_{ab}+V=-\sum_{a=2}^{N}\beta_{a}\sum_{b=1}^{N}e_{ab}+\exp{\left(\sum_{a=2}^{N}e_{1a}\right)}

@@

			\widehat{\Psi}(\bm{s})=\sum_{i=1}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{i},1}}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(L+2)}\prod_{l=1}^{L}\left(\lambda_{s_{l}}\left(2+L-l-\sum_{j=l}^{L}c_{j}\right)\right)^{c_{l}}\beta_{s_{l}}^{(1-c_{l})(1-\delta_{s_{l},1})}

@@

	\left(T^{T}\otimes T^{T}\right)_{s_{1},s_{2}}=e_{s_{1},s_{2}}^{s_{1},s_{2}}+(1-\delta_{s_{2},1})\beta_{s_{2}}e_{s_{1},s_{2}}^{s_{1},1}+(1-\delta_{s_{1},1})\beta_{s_{1}}e_{s_{1},s_{2}}^{1,s_{2}}+(1-\delta_{s_{1},1})(1-\delta_{s_{2},1})\beta_{s_{1}}\beta_{s_{2}}e_{s_{1},s_{2}}^{1,1}

@@

  \left(1+\sum_{j=i}^{L}\delta_{s_{j},1}\right) =\left(2+L-i-\sum_{j=i}^{L}(1-\delta_{s_{i},1})\right)

@@

	\widetilde{\Psi}(\bm{s})=\frac{\Gamma(2+\sum_{i=1}^{L}\delta_{s_{i},1})}{\Gamma(L+2)}\prod_{i=1}^{L}\left(\lambda_{s_{i}}\left(2+L-i-\sum_{j=i}^{L}(1-\delta_{s_{i},1})\right)\right)^{1-\delta_{s_{i},1}}

@@

		\widehat{\Psi}(\mathbf{s})=\sum_{c_{1}=0}^{1-\delta_{s_{1},1}}\ldots\sum_{c_{L}=0}^{1-\delta_{s_{L},1}}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(L+2)}\prod_{i=1}^{L}\left(\lambda_{s_{i}}\left(2+L-i-\sum_{j=i}^{L}c_{j}\right)\right)^{c_{i}}\left(\beta_{s_{i}}\right)^{(1-c_{i})(1-\delta_{s_{i},1})}
Indeed, when $c_{i}=0$ this formula replaces the contribution in the square brackets with a $\beta_{s_{i}}$. 

@@

        \Psi(\bm{s})&=\sum_{i=1}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{i},1}}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(L+2)}\prod_{l=1}^{L}\left(\lambda_{s_{l}}\left(2+L-l-\sum_{j=l}^{L}c_{j}\right)\right)^{c_{l}}\beta_{s_{l}}^{(1-c_{l})(1-\delta_{s_{l},1})}

@@

        \sum_{r=1}^{\ell}\sum_{d_{q_{r}}=2}^{N}
\sum_{i=1\,:\, i\neq q_{1},\ldots,q_{l}}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{i},1}}\sum_{p=1}^{\ell}\sum_{c_{q_{p}}=0}^{1}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(2+L)}
  \prod_{i=1\,:\, i\neq q_{1},\ldots,q_{\ell}}^{L}\left(\lambda_{s_{i}}\left(2+L-i-\sum_{j=i}^{L}c_{j}\right)\right)^{c_{i}}

@@

 \beta_{s_{i}}^{(1-c_{i})(1-\delta_{s_{i},1})}
 \prod_{t=1}^{\ell}\lambda_{d_{q_{t}}}\left(2+L-q_{t}-\sum_{j=q_{t}}^{L}c_{j}\right)^{c_{q_{t}}}\beta_{d_{q_{t}}}^{(1-c_{q_{t}})}

@@

  &\widehat{\Psi}(\ldots,a_{i},\ldots,d_{q_{r}},\ldots)=\sum_{i=1\,:\, i\neq q_{1},\ldots,q_{l}}^{L}\sum_{c_{i}=0}^{1-\delta_{s_{i},1}}\sum_{p=1}^{\ell}\sum_{c_{q_{p}}=0}^{1}\frac{\Gamma(2+L-\sum_{k=1}^{L}c_{k})}{\Gamma(2+L)}

@@

\prod_{i=1\,:\, i\neq q_{1},\ldots,q_{\ell}}^{L}\left(\lambda_{s_{i}}\left(2+L-i-\sum_{j=i}^{L}c_{j}\right)\right)^{c_{i}}\beta_{s_{i}}^{(1-c_{i})(1-\delta_{s_{i},1})}
  \prod_{t=1}^{\ell}\lambda_{d_{q_{t}}}\left(2+L-q_{t}-\sum_{j=q_{t}}^{L}c_{j}\right)^{c_{q_{t}}}\beta_{d_{q_{t}}}^{(1-c_{q_{t}})}
indeed, in coordinates $q_{r}$ with $r=1,\ldots \ell$ the species $1$ is replaced by all other possible species from $2$ to $N$.  Finally, considering all possible occupations $\bm{s}=(s_{1},\ldots,s_{L})$ we obtain \eqref{steadyStateH}. 

@@

	\widehat{\Psi}_{\bm{y}}(s_{1},\ldots,s_{m})=\sum_{i=1}^{L}\sum_{c_{i}=0}^{1}\frac{\Gamma(2+L-\sum_{k=1}^{m}c_{k})}{\Gamma(2+L)}\prod_{k=1}^{m}\lambda_{s_{k}}^{c_{k}}\beta_{s_{k}}^{1-c_{k}}\left(2+L-y_{k}-\sum_{j=k}^{m}c_{j}\right)^{c_{k}}
